---
title: "Health & Access to Care Dashboard"
author: Charles Jacobs
format: 
  dashboard:
    nav-buttons: [github]
    github: https://github.com/mine-cetinkaya-rundel/ld-dashboard
theme: lumen
fig-width: 10
fig-asp: 0.3
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(Hmisc)
library(gtsummary)
library(nhanesA)
library(scales)
library(DT)
library(gt)
library(broom)
theme_set(theme_minimal(base_size = 24, base_family = "sans"))
```

```{r}
#| label: load-data
#| message: false

hospital_raw <- readRDS("data/P_HUQ.rds")
body_raw <- readRDS("data/P_BMX.rds")
demo_raw <- readRDS("data/P_DEMO.rds")
```

```{r}
#| label: prep-data
body_cleaned <- body_raw %>%
  select(SEQN, BMXBMI)

demo_cleaned <- demo_raw %>%
  select(SEQN, RIDAGEYR, RIAGENDR, RIDRETH3, DMDEDUC2, INDFMPIR, RIDSTATR)

hospital_cleaned <- hospital_raw %>%
  select(SEQN, HUQ010, HUQ090, HUQ071, HUQ030)


merged_data <- demo_cleaned %>%
  left_join(body_cleaned, by = "SEQN") %>%
  left_join(hospital_cleaned, by = "SEQN") %>%
  filter(RIDSTATR == "Both interviewed and MEC examined") %>%
  filter(between(RIDAGEYR, 21, 79)) %>%
  filter(complete.cases(.)) %>%
  filter(HUQ071 != "Don't know") %>%
  filter(!HUQ010 %in% c("Don't know", "Refused")) %>%
  filter(!HUQ030 %in% c("Don't know", "Refused")) %>%
  mutate(
    HUQ030 = fct_recode(HUQ030, "Yes" = "There is more than one place"),
    HUQ030 = fct_relevel(HUQ030, "Yes", "No"),
    HUQ071 = fct_relevel(HUQ071, "No", "Yes"),
    HUQ090 = fct_relevel(HUQ090, "No", "Yes")
  ) %>%
  droplevels()
```

#  {.sidebar}

This dashboard displays statistics for:

|                 |                                     |
|-----------------|-------------------------------------|
| **Dataset**     | NHANES 2017-March 2020 Pre-pandemic |
| **Population**  | Adults ages 21 to 79                |
| **Sample size** | `{r} nrow(merged_data)`             |

------------------------------------------------------------------------

Variables featured in this dashboard:

|                           |            |
|---------------------------|------------|
| **BMI**                   | `BMXBMI`   |
| **Hospital stay**         | `HUQ071`   |
| **Overall health**        | `HUQ010`   |
| **Routine care location** | `HUQ030`   |
| **Ethnicity**             | `RIDRETH3` |
| **Mental health visit**   | `HUQ090`   |

------------------------------------------------------------------------

::: {.callout-note collapse="true"}
## Disclaimer

This dashboard summarizes cleaned NHANES data restricted to adults ages 21 to 79 with complete data on selected variables.
:::

# All

```{r}
#| label: all-values
#| results: hide

n_adults <- nrow(merged_data)

mean_bmi <- mean(merged_data$BMXBMI, na.rm = TRUE)

p_hosp_yes <- merged_data %>%
  count(HUQ071) %>%
  mutate(p = n / sum(n)) %>%
  filter(HUQ071 == "Yes") %>%
  pull(p)

p_mh_yes <- merged_data %>%
  count(HUQ090) %>%
  mutate(p = n / sum(n)) %>%
  filter(HUQ090 == "Yes") %>%
  pull(p)
```

## Row {height="20%"}

```{r}
#| content: valuebox
#| title: "Total adults"

list(
  icon = "people-fill",
  color = "primary",
  value = n_adults
)
```

```{r}
#| content: valuebox
#| title: "Mean BMI"

list(
  icon = "activity",
  color = "warning",
  value = round(mean_bmi, 1)
)
```

```{r}
#| content: valuebox
#| title: "Overnight Hospital stay in past year"

list(
  icon = "hospital",
  color = "danger",
  value = label_percent(accuracy = 0.1)(p_hosp_yes)
)
```

```{r}
#| content: valuebox
#| title: "Saw mental health professional in past year"

list(
  icon = "person-heart",
  color = "success",
  value = label_percent(accuracy = 0.1)(p_mh_yes)
)

```

## Row {height="40%"}

### Column {width="40%"}

```{r}
#| title: Overall health

merged_data %>%
  count(HUQ010) %>%
  mutate(p = n / sum(n)) %>%
  gt() %>%
  fmt_percent(
    columns = p,
    decimals = 1
  ) %>%
  cols_label(
    HUQ010 = "Overall health",
    n = "Count",
    p = "Proportion"
  )
```

### Column {width="60%"}

```{r}
#| title: BMI by overnight hospital stay

merged_data %>%
  ggplot(aes(x = HUQ071, y = BMXBMI, fill = HUQ071)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.15, alpha = 0.8) +
  scale_fill_manual(values = c("No" = "orangered", "Yes" = "royalblue")) +
  labs(
    x = "Overnight hospital stay in past year",
    y = "Body Mass Index (kg/m^2)")
   

```

## Row {height="40%"}

```{r}
#| title: Overnight hospital stay by mental health professional visit

ggplot(merged_data, aes(x = HUQ090, fill = HUQ071)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("No" = "orangered", "Yes" = "royalblue")) +
  scale_y_continuous(labels = label_percent()) +
  labs(
    x = "Seen mental health professional in past year",
    y = "Proportion",
    fill = "Overnight hospital stay"
  )
```

```{r}
#| title: Routine place for healthcare by overall health

ggplot(merged_data, aes(x = HUQ030, fill = HUQ010)) +
  geom_bar(position = "fill") +
  scale_fill_discrete(labels = c(
    "Excellent," = "Excellent",
    "Very good," = "Very Good",
    "Good," = "Good",
    "Fair, or" = "Fair",
    "Poor?" = "Poor"
  )) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = c(
    "There is no place" = "No",
    "Yes" = "Yes"
  )) +
  labs(
    title = "Routine Place for Healthcare by Overall Health Rating",
    x = "Routine Place for Healthcare",
    y = "Proportion",
    fill = "Overall Health Rating"
  ) +
  theme_minimal()
```

# BMI and Hospital Stay {orientation="columns"}

## Column {width="60%"}

```{r}
#| label: bmi-hosp-values
#| results: hide

hosp_yes <- merged_data %>% filter(HUQ071 == "Yes")
hosp_no  <- merged_data %>% filter(HUQ071 == "No")

mean_bmi_yes <- mean(hosp_yes$BMXBMI, na.rm = TRUE)
mean_bmi_no  <- mean(hosp_no$BMXBMI, na.rm = TRUE)

ttest_b <- t.test(BMXBMI ~ HUQ071, data = merged_data, conf.level = 0.90)
```

### Row {height="20%"}

```{r}
#| component: valuebox
#| title: "Mean BMI: Hospital stay = Yes"

list(
  icon = "clipboard2-pulse",
  color = "primary",
  value = round(mean_bmi_yes, 1)
)
```

```{r}
#| component: valuebox
#| title: "Mean BMI: Hospital stay = No"

list(
  icon = "clipboard2-pulse",
  color = "light",
  value = round(mean_bmi_no, 1)
)
```

### Row {height="40%"}

```{r}
#| component: valuebox
#| title: "Mean difference"

list(
  icon = "plus-slash-minus",
  color = "warning",
  value = round(unname(diff(rev(ttest_b$estimate))), 2)
)
```

### Row {.tabset height="40%"}

```{r}
#| title: BMI summary by overnight hospital stay

merged_data %>%
  group_by(HUQ071) %>%
  summarise(
    n = n(),
    Mean = mean(BMXBMI),
    SD = sd(BMXBMI),
    Median = median(BMXBMI),
    IQR = IQR(BMXBMI),
    Minimum = min(BMXBMI),
    Maximum = max(BMXBMI)
  ) %>%
  gt()
```

```{r}
#| title: BMI distribution

merged_data %>%
  ggplot(aes(x = HUQ071, y = BMXBMI, fill = HUQ071)) +
  geom_violin(alpha = 0.5, trim = FALSE) +
  geom_boxplot(width = 0.15, alpha = 0.8) +
  scale_fill_manual(values = c("No" = "orangered", "Yes" = "royalblue")) +
  labs(
    x = "Overnight hospital stay",
    y = "BMI",
    fill = NULL
  ) 

```

```{r}
#| title: t test results

tibble(
  estimate_yes = unname(ttest_b$estimate[1]),
  estimate_no = unname(ttest_b$estimate[2]),
  difference = unname(diff(rev(ttest_b$estimate))),
  conf_low = ttest_b$conf.int[1],
  conf_high = ttest_b$conf.int[2],
  p_value = ttest_b$p.value
) %>%
  gt()
```

## Column {width="40%"}

```{r}
#| title: Data

merged_data %>%
  select(SEQN, BMXBMI, HUQ071) %>%
  arrange(SEQN) %>%
  datatable(
    colnames = c("ID", "BMI", "Overnight hospital stay"),
    options = list(
      dom = "lrtip",
      paging = TRUE,
      lengthMenu = c(5, 10, 25),
      pageLength = 10,
      scrollY = TRUE
    )
  )
```

# Ethnicity and BMI {orientation="columns"}

## Column {width="60%"}

```{r}
#| label: ethnicity-values
#| results: hide

anova_C <- aov(BMXBMI ~ RIDRETH3, data = merged_data)
anova_p <- summary(anova_C)[[1]][["Pr(>F)"]][1]
```

### Row {height="20%"}

```{r}
#| component: valuebox
#| title: "Ethnic groups"

list(
  icon = "people",
  color = "primary",
  value = n_distinct(merged_data$RIDRETH3)
)
```

```{r}
#| component: valuebox
#| title: "Overall mean BMI"

list(
  icon = "activity",
  color = "warning",
  value = round(mean(merged_data$BMXBMI), 1)
)
```

```{r}
#| component: valuebox
#| title: "ANOVA p-value"

list(
  icon = "bar-chart-line",
  color = "danger",
  value = signif(anova_p, 3)
)
```

### Row {height="40%"}

```{r}
#| title: BMI by ethnicity

merged_data %>%
  ggplot(aes(x = RIDRETH3, y = BMXBMI, fill = RIDRETH3)) +
  geom_boxplot(alpha = 0.7) +
  scale_x_discrete(labels = c(
    "Other Hispanic" = "Hispanic",
    "Non-Hispanic White" = "White",
    "Non-Hispanic Black" = "Black",
    "Non-Hispanic Asian" = "Asian",
    "Other Race - Including Multi-Racial" = "Other"
  )) +
  labs(
    x = "Ethnicity",
    y = "BMI",
    fill = NULL
  ) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1),
        legend.position = "none")
```

### Row {.tabset height="40%"}

```{r}
#| title: Maternal age and parity

merged_data %>%
  group_by(RIDRETH3) %>%
  summarise(
    n = n(),
    Mean = mean(BMXBMI),
    SD = sd(BMXBMI),
    Median = median(BMXBMI),
    IQR = IQR(BMXBMI),
    Minimum = min(BMXBMI),
    Maximum = max(BMXBMI)
  ) %>%
  gt()
```

```{r}
#| title: Tukey HSD

broom::tidy(TukeyHSD(anova_C, conf.level = 0.90)$RIDRETH3) %>%
  gt()
```

## Column {width="40%"}

```{r}
#| title: Data

merged_data %>%
  select(SEQN, BMXBMI, RIDRETH3) %>%
  arrange(SEQN) %>%
  datatable(
    colnames = c("ID", "BMI", "Ethnicity"),
    options = list(
      dom = "lrtip",
      paging = TRUE,
      lengthMenu = c(5, 10, 25),
      pageLength = 10,
      scrollY = TRUE
    )
  )
```

# Associations 

## Column {width="60%"}
```{r}
#| label: assoc-values
#| results: hide

risk_table <- table(
  MentalHealthProfessional = merged_data$HUQ090,
  OvernightHospital = merged_data$HUQ071
)

a <- risk_table["Yes", "Yes"]
b <- risk_table["Yes", "No"]
c <- risk_table["No", "Yes"]
d <- risk_table["No", "No"]

risk_yes <- a / (a + b)
risk_no  <- c / (c + d)

risk_difference <- risk_yes - risk_no
relative_risk   <- risk_yes / risk_no
odds_ratio      <- (a * d) / (b * c)

jk_table <- table(merged_data$HUQ030, merged_data$HUQ010)
chisq_e <- chisq.test(jk_table)
```


### Row {height="20%"}

```{r}
#| component: valuebox
#| title: "Relative risk"

list(
  icon = "arrow-left-right",
  color = "primary",
  value = round(relative_risk, 2)
)

```

```{r}
#| component: valuebox
#| title: "Odds ratio"

list(
  icon = "percent",
  color = "warning",
  value = round(odds_ratio, 2)
)
```


```{r}
#| component: valuebox
#| title: "Chi-square p-value"

list(
  icon = "grid-3x3-gap",
  color = "danger",
  value = signif(chisq_e$p.value, 3)
)
```

### Row {height="40%"}

```{r}
#| title: Hospital stay by mental health visit

ggplot(merged_data, aes(x = HUQ090, fill = HUQ071)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("No" = "orangered", "Yes" = "royalblue")) +
  scale_y_continuous(labels = label_percent()) +
  labs(
    x = "Seen mental health professional",
    y = "Proportion",
    fill = "Overnight hospital\nstay"
  )
```

### Row {.tabset height="40%"}
```{r}
#| title: Risk estimates

tibble(
  risk_mh_yes = risk_yes,
  risk_mh_no = risk_no,
  risk_difference = risk_difference,
  relative_risk = relative_risk,
  odds_ratio = odds_ratio
) %>%
  gt()
```

```{r}
#| title: Routine care and overall health

ggplot(merged_data, aes(x = HUQ030, fill = HUQ010)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = label_percent()) +
  labs(
    x = "Routine place for healthcare",
    y = "Proportion",
    fill = "Overall health"
  )
```

```{r}
#| title: Chi-square test

tibble(
  statistic = unname(chisq_e$statistic),
  df = unname(chisq_e$parameter),
  p_value = chisq_e$p.value
) %>%
  gt()
```

## Column {width="40%"}

```{r}
#| title: Data

merged_data %>%
  select(SEQN, HUQ090, HUQ071, HUQ030, HUQ010) %>%
  arrange(SEQN) %>%
  datatable(
    colnames = c(
      "ID",
      "Mental health professional",
      "Overnight hospital stay",
      "Routine healthcare place",
      "Overall health"
    ),
    options = list(
      dom = "lrtip",
      paging = TRUE,
      lengthMenu = c(5, 10, 25),
      pageLength = 10,
      scrollY = TRUE
    )
  )
```

Data
```{r}
merged_data %>%
  select(SEQN, BMXBMI, HUQ010, HUQ090, HUQ071, HUQ030, RIDRETH3) %>%
  arrange(SEQN) %>%
  datatable(
    colnames = c(
      "ID",
      "BMI",
      "Overall health",
      "Mental health professional",
      "Overnight hospital stay",
      "Routine place for healthcare",
      "Ethnicity"
    ),
    options = list(dom = "ftp", paging = TRUE)
  )
```